### 2021-02-27 doc/article.tex
Rough outline of article latex document and IOP required statements

\documentclass[12pt]{iopart}

\usepackage{graphicx}
\graphicspath{{../fig/}}

\begin{document}

\title[PREPRINT by Doss-Gollin, Farnham, Lall, and Modi]{How unprecedented was the February 2021 Texas cold snap?}

\author{James Doss-Gollin$^1$, David J. Farnham$^2$, Upmanu Lall,$^3$ and Vijay Modi$^4$}
\address{$^1$ Department of Civil and Environmental Engineering, Rice University, Houston, TX}
\address{$^2$ Demo}
\address{$^3$ Demo}
\address{$^4$ Demo}
\ead{jdossgollin@rice.edu}
\vspace{10pt}

\begin{abstract}
  We need to write an abstract!
\end{abstract}
\noindent{\it Keywords: Climate Dynamics, Cold Extremes, Energy Demand, Texas}

%\submitto{\ERL}
\maketitle
\nocite{*}

\section{Introduction}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{historic_events_era5.pdf}
  \caption{CAPTION}\label{fig:historic_era5}
\end{figure}

\section{Methods}

\section{Results}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{local_rt_era5.pdf}
  \caption{CAPTION}\label{fig:local_era5}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{ERCOT_HDD_IDF_MLE_popweighted.pdf}
  \caption{CAPTION}\label{fig:idf_weighted}
\end{figure}

\section{Discussion}

\section{Conclusions}

\clearpage
\section*{Acknowledgements}

The authors thank Adam Massmann and Vivek Srikrishnan for insightful comments on a preliminary draft of this manuscript.
The authors thank\ldots.

\bibliographystyle{unsrt}
\bibliography{references}

\clearpage
\section{Supplemental Results}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{historic_events_bk.pdf}
  \caption{CAPTION}\label{fig:historic_bk}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{historic_events_era5_TX.pdf}
  \caption{CAPTION}\label{fig:historic_tx}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{local_rt_ghcnd.pdf}
  \caption{CAPTION}\label{fig:local_ghcnd}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{ERCOT_HDD_IDF_MLE_unweighted.pdf}
  \caption{CAPTION}\label{fig:idf_unweighted}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{ERCOT_HDD_IDF_plotpos_popweighted.pdf}
  \caption{CAPTION}\label{fig:idf_lmoments_weighted}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{ERCOT_HDD_IDF_plotpos_unweighted.pdf}
  \caption{CAPTION}\label{fig:idf_lmoments_unweighted}
\end{figure}

\end{document}