### 2021-12-06 doc/agu21/at-a-glance.tex

\begin{block}{Human Impacts}
    \begin{itemize}
        \item Electricity demand would have surpassed the all-season record had generation kept up \cite{ercotpublic_outagesv1:2021}
        \item Over \SI{30000}{\mega\watt} of lost output (mostly natural gas; \cite{busby_cascadingrisks:2021}) in TX for $> \SI{2}{days}$
        \item Grid within minutes of catastrophic failure (\cref{fig:ercot-frequency})
        \item Over 100 people died \cite{mulcahy_urideath:2021}
        \item Estimated \$130 billion damages \cite{busby_cascadingrisks:2021}
        \item Electricity outages caused cascading failures of water supply and other critical infrastructure
        \item Marginalized communities disproportionately affected \cite{dobbins_blackoutdisparity:2021}
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{magness_frequency.pdf}
            \caption{
                As demand spiked and generation failed, the Texas grid came within minutes of catastrophic failure.
                Figure from ERCOT~\cite{magness_review:2021}.
            }
            \label{fig:ercot-frequency}
        \end{figure}
    \end{framed}
\end{block}