### 2021-12-06 doc/agu21/poster.tex

\documentclass[serif,mathserif,final]{beamer}
\mode<presentation>{\usetheme{Lankton}}

\usepackage[
  orientation=landscape,
  width=48in,height=36in,%size=a0, % paper size
  scale=1.25 % font scale factor
  ]{beamerposter}
 \setbeamertemplate{caption}[numbered] % number figs

% Graphics Path
\usepackage{graphicx}
\graphicspath{{../../fig/}{./}}

\usepackage[ttscale=0.8]{libertine} % better font

% OTHER PACKAGE CALLS
\usepackage[utf8]{inputenc}
\usepackage{
  amsmath,
  amsfonts,
  amssymb,
  xspace,
  framed,
  siunitx,
  nth,
  physics,
  nicefrac,
}
\sisetup{round-mode = figures, round-precision = 3}
\usepackage{cleveref}

%------ USE BIBLATEX WITH EXTREMELY MINAMALIST STYLE -----
% Use BibLaTeX
\usepackage[
	backend=biber,
	url=false,
	isbn=false,
	doi=false,
	maxnames = 2,
	style=numeric-comp,
	uniquename=false,
	uniquelist=false,
	firstinits=true,
	sorting=none,
	natbib=false,
	isbn=false,
]{biblatex}
\addbibresource{library.bib}
\renewbibmacro{in:}{}

% One-paragraph bibliography environment
\defbibenvironment{bibliography}
{\list%
  {\printtext[labelnumberwidth]{%
    \printfield{prefixnumber}%
    \printfield{labelnumber}}%
  \ifentrytype{article}{}{}}%
    {\setlength{\leftmargin}{0pt}%
    \setlength{\topsep}{0pt}}%
    \renewcommand*{\makelabel}[1]{##1}}%
  {\endlist}%
{\mkbibitem}

% \mkbibitem just prints item label and non-breakable space
\makeatletter
\newcommand{\mkbibitem}{\@itemlabel\addnbspace}
\makeatother

% Add breakable space between bibliography items
\renewcommand*{\finentrypunct}{\addperiod\space}

% Set fields to clear
\AtEveryBibitem{%
	\clearfield{month}%
	\clearfield{day}%
  \clearlist {language}%why this has to be clear list.....
  \clearfield{pages}%
	\clearfield{pagetotal}%
	\clearfield{eprinttype}%
	\clearfield{eprint}%
	\clearfield{number}%
	\clearfield{volume}%
	\clearfield{issue}%
	\clearfield{name:given}
	\clearfield{name:first}
	\ifentrytype{article}{%
		\clearfield{title}%
	}%
}

%------- GLOSSARIES PACKAGE -------
%\usepackage[acronym]{glossaries}
%\makeglossaries

%-- Header and footer information ----------------------------------
\newcommand{\footleft}{\textbf{Paper}: How unprecedented was the February 2021 Texas cold snap? \textit{Environmental Research Letters} (2021). \href{https://doi.org/10.1088/1748-9326/ac0278}{doi.org/10.1088/1748-9326/ac0278}.}
\newcommand{\footright}{\textbf{Codes}: \href{https://github.com/jdossgollin/2021-TXtreme/}{github.com/jdossgollin/2021-TXtreme/}}
\title{How Unprecedented Was the February 2021 Texas Cold Snap?}
\author{James Doss-Gollin\inst{*}\inst{1} \quad David J. Farnham\inst{2}\inst{3} \quad Vijay Modi\inst{4} \quad Upmanu Lall\inst{5}}
\institute{\inst{*}\href{mailto:jdossgollin@rice.edu}{jdossgollin@rice.edu} \quad \inst{1} Department of Civil and Environmental Engineering, Rice University \quad  \inst{2} ClimateAI \quad \inst{3} Department of Global Ecology, Carnegie Institution for Science \\ \inst{4} Department of Mechanical Engineering, Columbia University \quad \inst{5} Department of Earth and Environmental Engineering, Columbia University}

%-------------------------------------------------------------------

%-- Main Document --------------------------------------------------
\begin{document}
\begin{frame}{}
  \begin{columns}[t]

    \nocite{doss-gollin_txtreme:2021,doss-gollin_codes_txtreme:2021}

    %-- Column 1 ---------------------------------------------------
    \begin{column}{0.275\linewidth}

      \input{at-a-glance.tex}
      \input{key-findings.tex}
      \begin{block}{References}
        \renewcommand*{\bibfont}{\footnotesize}
        \printbibliography[heading=none]
      \end{block}

    \end{column}%1

    %-- Column 2 ---------------------------------------------------
    \begin{column}{0.325\linewidth}

      \input{how-cold.tex}
      \input{inferred-heating.tex}

    \end{column}

    %-- Column 3 ---------------------------------------------------
    \begin{column}{0.325\linewidth}

      \input{aggregate_rt.tex}
      \input{spatial_rt.tex}

    \end{column}

  \end{columns}
\end{frame}
\end{document}