### 2021-12-06 doc/agu21/how-cold.tex

\begin{block}{Historic Cold Snaps}
    Visualizing temperature anomalies facilitates identification of large-scale weather patterns superimposed on long-term climatological averages.
    We compare the February 2021 cold snap (bottom row) to four historic cold snaps.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth, trim = 0.25in 0 0 0.75in, clip]{historic_events_era5.pdf}
            \caption{
                \textbf{Previous cold snaps demonstrate a qualitative precedent for severe cold.}
                Data shows anomalies of daily mean temperatures from the ERA-5 reanalysis \cite{hersbach_era5:2020}.
            }\label{fig:historic_era5}
        \end{figure}
    \end{framed}
\end{block}